\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue!50!black,citecolor=blue!50!black,urlcolor=blue!50!black}
\usepackage{listings}
\lstset{basicstyle=\ttfamily\small,breaklines=true,columns=fullflexible}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit,backgrounds}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{array}
\usepackage{titlesec}
\titlespacing*{\section}{0pt}{1.4em}{0.7em}
\titlespacing*{\subsection}{0pt}{1.1em}{0.5em}
\setlength{\parskip}{0.4em}
\setlength{\parindent}{0pt}

\title{\textbf{TPM-Attest: Hardware-Rooted Integrity Attestation as a Kernel-Level Anti-Cheat Alternative for Linux}}

\author[1]{Anudeep Gedela}
\author[1,*]{Dr.\ A.\ Yaswanth}
\affil[1]{GITAM School of Technology, GITAM (Deemed to be University), Visakhapatnam, India}
\affil[*]{Corresponding author -- Assistant Professor, Department of Computer Science \& Engineering, \texttt{yamanapu@gitam.edu}}
\date{}

\begin{document}
\maketitle

\begin{abstract}
Multiplayer PC gaming on Linux faces a structural problem: the anti-cheat systems that publishers require operate as proprietary Ring~0 kernel modules that are architecturally incompatible with Linux's security model, GPL licensing, and stable ABI guarantees. We argue the right response is not to port these invasive modules to Linux, but to replace them entirely. This paper presents \textbf{TPM-Attest}, a hardware-rooted remote attestation framework that uses the Trusted Platform Module (TPM)~2.0 and the Linux Integrity Measurement Architecture (IMA) to prove, cryptographically, that a client booted cleanly and ran only authorised software -- without any kernel driver, without proprietary code, and without scanning player memory. The system intercepts Epic Online Services (EOS) SDK calls via a userspace \texttt{LD\_PRELOAD} hook, gates session access on a live TPM quote bound to a server-issued nonce, and constructs an index-prefixed Merkle tree over the IMA log that is immune to duplicate-leaf collision attacks. Across 500 constructed tamper sessions we achieve a 100\% detection rate; incremental leaf caching reduces repeat-attestation latency to under 3~seconds on real TPM~2.0 hardware. A controlled red-team evaluation against a live demo game confirms all four file-backed attack vectors are blocked while precisely characterising the two confirmed bypass conditions. The full implementation is released as open-source software.
\end{abstract}

\section{Introduction}
\label{sec:intro-ref}

Valve's Proton compatibility layer has made thousands of Windows titles playable on Linux~\cite{valve2018proton}, yet competitive multiplayer games remain off-limits for most Linux players. The barrier is not the game logic -- it is the anti-cheat. Systems like Easy Anti-Cheat and BattlEye rely on Ring~0 kernel drivers that assume a Windows environment, conflict with Linux's stable ABI guarantees, and cannot be shipped under GPL. The result is straightforward: Linux players are locked out.

Worse, this lockout carries a hidden cost for everyone. A recent audit of four widely deployed kernel-mode anti-cheat systems found that two behaved indistinguishably from rootkits -- broad OS visibility, covert channels, and access far exceeding any legitimate integrity-checking purpose~\cite{dorner2024rootkit}. Players on Windows accept these risks because they have no choice; Linux's kernel rejects such code, which is why Linux support is absent.

We take a different approach. Instead of scanning memory at runtime, TPM-Attest asks one question at session startup: \textit{did this machine boot into a known-good state and execute only authorised software?} The TPM~2.0 microcontroller, present on virtually every modern motherboard, can answer this with a hardware-signed cryptographic quote~\cite{tcg2018}. Paired with the Linux Integrity Measurement Architecture -- a kernel subsystem maintaining a tamper-evident log of every executed binary and loaded library since boot~\cite{sailer2004ima} -- the result is an attestation report the server can verify in under 15 seconds. A clean chain permits the session; any deviation denies access before the game reaches the network~\cite{rats2023}.

The key contributions of this paper are:
\begin{enumerate}[nosep]
  \item An open-source, end-to-end TPM-Attest framework combining TPM~2.0 quotes and IMA measurement logs for Linux gaming integrity verification.
  \item An index-prefixed Merkle tree construction immune to duplicate-leaf collision attacks (CVE-2012-2459 class).
  \item A dynamic \texttt{LD\_PRELOAD} interception hook gating EOS SDK sessions on real-time attestation results.
  \item A comprehensive experimental evaluation including latency profiling, tamper-detection rates, and a comparative analysis against Keylime and kernel-mode anti-cheat baselines.
  \item \textbf{VOID SECTOR} -- a complete playable 2-D space-shooter demo game whose session launch is gated end-to-end by the TPM-Attest pipeline, providing a live, reproducible demonstration of the full attestation flow (game client $\to$ hook $\to$ daemon $\to$ TPM $\to$ server $\to$ session granted/denied).
\end{enumerate}

\section{Background and Foundations}

\subsection{TPM 2.0 and Platform Configuration Registers (PCRs)}
The TPM~2.0 microcontroller operates as a secure cryptoprocessor physically isolated from the host CPU~\cite{tcg2018}. Platform Configuration Registers (PCRs) are volatile memory slots storing hashes representing the platform's boot and software state. PCRs are updated via a one-way \textbf{extend} operation:

\begin{equation}
  \text{PCR}_{\text{new}} = \text{SHA-256}\!\left(\text{PCR}_{\text{old}} \,\|\, m\right)
  \label{eq:pcr-extend}
\end{equation}

\noindent where $m$ is the new measurement value. This one-way chaining prevents register rollback or state forging. A TPM quote (\texttt{tpm2\_quote}) extracts a signed digest of selected PCRs using an asymmetric Attestation Key (AK), incorporating a caller-supplied challenge nonce $\eta$ to guarantee freshness~\cite{tpm2tools2024}.

\subsection{Integrity Measurement Architecture (IMA)}
Linux IMA is a kernel subsystem that intercepts system calls mapping files for execution (e.g., \texttt{execve}, \texttt{mmap})~\cite{lko2021}. It computes file hashes, appends them to an in-memory ASCII log, and extends the measurements into PCR~10. The log is anchored by the \texttt{boot\_aggregate} measurement (representing PCRs 0--7), linking runtime executions to the physical boot state. The aggregate is defined as:

\begin{equation}
  \texttt{boot\_aggregate} = \text{SHA-256}\!\left(\bigoplus_{i=0}^{7} \text{PCR}_i\right)
  \label{eq:boot-agg}
\end{equation}

\subsection{Shared Library Interception (\texttt{LD\_PRELOAD})}
On Unix systems, the dynamic linker allows loading custom shared libraries before standard dependencies via the \texttt{LD\_PRELOAD} environment variable~\cite{kerrisk2010}. By exporting matching function symbols (such as those in the Epic Online Services SDK), we intercept SDK network calls at runtime, evaluate local attestation status, and block access transparently.

\subsection{Index-Prefixed Merkle Tree}
Standard binary Merkle trees are susceptible to duplicate-leaf collision attacks~\cite{bitcoin2012cve}. TPM-Attest prevents this by prepending the absolute log index $i$ to the leaf hash preimage:

\begin{equation}
  \ell_i = \text{SHA-256}\!\left(i \,\|\, \text{pcr}_i \,\|\, \text{tmpl\_hash}_i \,\|\, \text{file\_hash}_i \,\|\, \text{filename}_i\right)
  \label{eq:leaf}
\end{equation}

Internal nodes are computed as:
\begin{equation}
  n_{\text{parent}} = \text{SHA-256}\!\left(\ell_{\text{left}} \,\|\, \ell_{\text{right}}\right)
  \label{eq:internal}
\end{equation}

Since $\ell_i$ is bound to its absolute position $i$, duplication or reordering changes the leaf hash and is rejected by the verifier.


\section{Related Work}
\label{sec:related-work}

The idea of using hardware measurements to verify software integrity on a running host dates to Sailer et al., who introduced IMA and showed that a remote verifier can reliably detect compromised binaries without modifying the client kernel~\cite{sailer2004ima}. \textbf{Keylime}, developed at MIT Lincoln Laboratory and now a CNCF-graduated project, industrialised this into a scalable agent/registrar/verifier architecture for cloud, edge, and IoT deployments~\cite{keylime}. TPM-Attest draws on the same measurement primitives but targets a different problem: gating a single consumer desktop's game session at the moment an EOS SDK call fires, rather than monitoring a server fleet continuously.

The strongest motivation for this work is Dorner and Klausner's 2024 independent audit of four widely deployed kernel-level anti-cheat systems, two of which displayed rootkit-like characteristics -- unrestricted OS visibility, covert communication channels, and access patterns inconsistent with their stated purpose~\cite{dorner2024rootkit}. TPM-Attest is a direct architectural response: the trust anchor moves into hardware and the kernel driver disappears entirely. Alangari and Alharbi's 2025 survey of anti-cheat defences confirms that no existing Linux-native solution successfully combines hardware-rooted trust with EOS SDK integration and GPL compatibility~\cite{alangari2025review}. Coker et al.\ formalised the properties a remote attestation protocol must satisfy~\cite{coker2011principles}; our nonce challenge-response design satisfies their freshness and binding requirements directly. Bratus et al.\ and Petroni et al.\ both argue that passive load-time measurement and active runtime scanning are complementary rather than competing~\cite{bratus2008,petroni2004} -- a position our Section~8 deployment recommendations adopt explicitly.


\section{Threat Model and Security Goals}
\label{sec:threat-model}

\subsection{Adversary Capabilities}
We model a motivated adversary who has obtained root access to the target machine and seeks to gain an unfair advantage in a multiplayer game without detection. The adversary may replace or patch game binaries on disk, inject cheat libraries at process-launch time, load unsigned kernel modules, capture and replay a previously accepted attestation report, or run the game inside a virtual machine equipped with a software-emulated TPM.

\subsection{Prevented Attacks}
\begin{enumerate}[nosep]
  \item \textbf{Binary and library tampering:} Any modification to an executed file is recorded by IMA and changes the Merkle root, causing the server to reject the report.
  \item \textbf{Kernel modification:} Changes to the kernel image extend PCR~9 with a new measurement, breaking the stored baseline.
  \item \textbf{Report replay:} Every report is cryptographically bound to a single-use server nonce; captured reports cannot be reused.
  \item \textbf{TPM emulation (vTPM/swtpm):} Software TPMs lack a manufacturer-signed Endorsement Key certificate; the server rejects them under \texttt{STRICT\_EK\_VERIFICATION} mode.
  \item \textbf{Unauthorised MOK kernel:} A custom kernel enrolled via Machine Owner Key alters PCR~7; the server detects the deviation and denies the session.
\end{enumerate}

\subsection{Out-of-Scope Attacks}
TPM-Attest makes no claims about physical bus-level attacks (SPI/LPC sniffing, PCIe DMA) that bypass the OS entirely. Runtime code injection via \texttt{ptrace} or anonymous executable mappings into already-measured processes also falls outside scope: IMA measures files at load time, not memory pages written after the fact. Section~\ref{sec:redteam} confirms these bypass conditions experimentally; Section~\ref{sec:limitations} discusses mitigations.


\section{System Design and Architecture}
\label{sec:design}

TPM-Attest is built around three loosely coupled components that together form the complete attestation pipeline (Figure~\ref{fig:architecture}). Each component communicates only through well-defined interfaces -- Unix socket or HTTP -- so any layer can be replaced or audited independently without touching the others.

\begin{figure}[h]
\centering
\begin{tikzpicture}[
    box/.style={draw,rectangle,rounded corners,align=center,minimum height=0.9cm,font=\small},
    >=Stealth
]
\node[box,minimum width=3.0cm] (client) {Game Client};
\node[box,below=0.9cm of client,minimum width=3.0cm] (hook) {\texttt{eac\_hook.so}\\(LD\_PRELOAD Hook)};
\node[box,right=2.0cm of hook,minimum width=3.0cm] (shim) {\texttt{shim.py}\\(Local Daemon)};
\node[box,below=0.9cm of shim,minimum width=3.0cm] (report) {\texttt{report.py}\\(Report Generator)};
\node[box,below left=1.1cm and -0.2cm of report,minimum width=2.2cm] (tpm) {TPM 2.0\\Hardware};
\node[box,below right=1.1cm and -0.2cm of report,minimum width=2.2cm] (ima) {IMA\\Security};
\draw[->] (client) -- node[right,font=\tiny]{calls EOS SDK APIs} (hook);
\draw[<->] (hook) -- node[above,font=\tiny]{Unix socket} (shim);
\draw[->] (shim) -- node[right,font=\tiny]{queries agent} (report);
\draw[->] (report) -- node[midway,left,font=\tiny]{reads PCR} (tpm);
\draw[->] (report) -- node[midway,right,font=\tiny]{reads IMA log} (ima);
\begin{scope}[on background layer]
\node[draw,dashed,inner sep=0.5cm,fit=(client)(hook)(shim)(report)(tpm)(ima),
      label={[font=\small]above:\textbf{Attested Machine}}] (machine) {};
\end{scope}
\node[box,below=1.7cm of machine.south,minimum width=4.6cm] (server) {Verification Server\\\texttt{server/main.py}};
\draw[->] (machine.south) -- node[midway,right,font=\tiny]{HTTP POST \texttt{/attest}} (server);
\end{tikzpicture}
\caption{TPM-Attest system architecture: client-side attestation agent, local interception hook/daemon, and remote verification server.}
\label{fig:architecture}
\end{figure}

\subsection{Phase 1: Client-Side Attestation Agent}
\begin{itemize}[nosep]
  \item \textbf{PCR Reader (\texttt{agent/pcr\_reader.py}):} Spawns \texttt{tpm2\_pcrread} to retrieve SHA-256 PCR bank for indices $\{0,1,4,7,9,10\}$.
  \item \textbf{IMA Reader (\texttt{agent/ima\_reader.py}):} Parses the kernel measurement log and computes the Merkle tree per Equations~\ref{eq:leaf}--\ref{eq:internal}.
  \item \textbf{Report Generator (\texttt{agent/report.py}):} Invokes \texttt{tpm2\_quote} under a secure temp directory (permissions \texttt{0o700}) signing PCRs with the Attestation Key (AK) at handle \texttt{0x81000000}, incorporating the server-issued nonce $\eta$.
\end{itemize}

\subsection{Phase 2: Local Interception Hook and Daemon}
\begin{itemize}[nosep]
  \item \textbf{Dynamic Hook (\texttt{phase3/eac\_hook.c}):} Intercepts \texttt{EOS\_AntiCheatClient\_AddNotifyMessageToServer}. Establishes a blocking Unix-socket connection with a 30-second \texttt{SO\_RCVTIMEO} timeout. Uses a custom fail-closed \texttt{parse\_json\_bool} parser to prevent string-injection bypasses.
  \item \textbf{Local Daemon (\texttt{phase3/shim.py}):} Fetches a fresh nonce, requests the agent to generate the report, transmits it to \texttt{/attest}, and returns the result over the socket.
\end{itemize}

\subsection{Phase 3: Verification Server}
The verification server (\texttt{server/main.py}) is a stateless FastAPI application backed by a lightweight SQLite database. It stores enrolled AK public keys, EK certificate records, consumed nonces, and pinned PCR baselines per machine. Keeping all trust state server-side -- never on the client -- ensures a compromised machine cannot alter its own records. The nonce lifecycle is strictly single-use (Equation~\ref{eq:nonce}): a nonce is created on \texttt{GET /challenge}, consumed atomically on \texttt{POST /attest}, and never reissued.

\begin{equation}
  \eta \leftarrow \text{CSPRNG}(32\text{ bytes}), \quad \eta \notin \mathcal{N}_{\text{used}}
  \label{eq:nonce}
\end{equation}

\section{Security Protocols and Hardening}

\subsection{Nonce Challenge-Response Protocol}
The nonce protocol is the first line of defence against report replay. The client calls \texttt{GET /challenge} to receive a fresh $\eta$; the server timestamps it and records it as pending. When the report arrives at \texttt{POST /attest}, the server verifies the TPM quote signature, confirms that $\eta$ appears verbatim inside the quote's qualifying data, then deletes $\eta$ -- all within a single database transaction. To further bound the attack window, the server additionally enforces a maximum nonce age of 30~seconds (Equation~\ref{eq:freshness}):

\begin{equation}
  |\,t_{\text{now}} - t_{\eta}\,| \leq \Delta t_{\max}, \quad \eta \notin \mathcal{N}_{\text{used}}
  \label{eq:freshness}
\end{equation}

\noindent Both conditions must hold simultaneously. The 30-second bound is reinforced client-side via the \texttt{SO\_RCVTIMEO} socket timeout in \texttt{shim.py}, so a slow or blocked daemon does not leave a stale nonce window open.

\subsection{Dynamic Policy Enforcement}
\begin{enumerate}[nosep]
  \item \textbf{\texttt{STRICT\_EK\_VERIFICATION}:} Requires a valid X.509 EK certificate; raw public keys are rejected, blocking \texttt{swtpm} register-spoofing.
  \item \textbf{\texttt{REQUIRE\_PCR7\_PINNING}:} Validates PCR~7 against the baseline, ensuring Secure Boot is active and rejecting unauthorised MOK kernels.
  \item \textbf{\texttt{REQUIRE\_IMA\_MINIMUM\_ENTRIES}:} Validates that the client submits a minimum number of IMA measurements, preventing an attacker from simply disabling the IMA log post-boot.
\end{enumerate}

\section{Experimental Results}
\label{sec:evaluation}

All experiments ran on a consumer desktop: Intel Core i7-10700, 16~GB RAM, Ubuntu~22.04 LTS with kernel~6.5, Secure Boot enabled, and an Intel PTT firmware TPM~2.0. We chose a firmware TPM deliberately -- it is by far the most common configuration in consumer gaming PCs, and any result obtained on a discrete chip would be unrepresentative of the actual deployment environment. No special kernel patches or elevated privileges beyond normal \texttt{tpm2-tools} device-node access were required.

\subsection{Tamper Detection Accuracy}

We scripted 500 attestation sessions -- 100 per tamper category -- each using a fresh server-issued nonce. For every session we applied the tamper condition, triggered attestation, and recorded whether the server rejected the report before issuing a token. Table~\ref{tab:detection} summarises the results.

\begin{table}[h]
\centering
\caption{Tamper detection results across 500 attestation sessions.}
\label{tab:detection}
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Tamper Scenario} & \textbf{Sessions} & \textbf{Detected} & \textbf{Missed} & \textbf{Detection Rate} \\
\midrule
Binary Substitution          & 100 & 100 & 0 & 100.0\% \\
IMA Log Truncation           & 100 & 100 & 0 & 100.0\% \\
IMA Log Duplication          & 100 & 100 & 0 & 100.0\% \\
Nonce Replay                 & 100 & 100 & 0 & 100.0\% \\
vTPM / swtpm Emulation       & 100 & 100 & 0 & 100.0\% \\
\midrule
\textbf{Total}               & 500 & 500 & 0 & \textbf{100.0\%} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Latency Profiling}

We measured wall-clock time for each pipeline stage over 50 independent runs, with the IMA log at approximately 8,000 entries -- a realistic size for a desktop that has been running for a few hours. Table~\ref{tab:latency} reports means and standard deviations.

\begin{table}[h]
\centering
\caption{Attestation pipeline latency (averaged over 50 runs, Intel PTT).}
\label{tab:latency}
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Pipeline Stage} & \textbf{Mean (s)} & \textbf{Std Dev (s)} \\
\midrule
IMA Log Read \& Merkle Build (first run, $\sim$8k entries) & 12.20 & 0.43 \\
IMA Log Read \& Merkle Build (incremental/cached)          &  0.18 & 0.02 \\
TPM Quote Generation (\texttt{tpm2\_quote})                &  2.10 & 0.11 \\
Network Round-Trip + Server Verification                    &  0.70 & 0.08 \\
\midrule
\textbf{Total -- First Attestation}                        & 15.00 & 0.52 \\
\textbf{Total -- Repeat Attestation (cached)}              &  2.98 & 0.15 \\
\bottomrule
\end{tabular}
\end{table}

The total attestation latency $T_{\text{total}}$ is modeled as the sum of three independent pipeline stages:

\begin{equation}
  T_{\text{total}} = T_{\text{IMA}} + T_{\text{quote}} + T_{\text{net}}
  \label{eq:latency}
\end{equation}

\noindent where $T_{\text{IMA}}$ is the IMA log read and Merkle build time, $T_{\text{quote}}$ is the TPM quote generation time, and $T_{\text{net}}$ is the network round-trip and server verification time. With incremental leaf caching (Section~\ref{sec:design}), the repeat-attestation model reduces to:

\begin{equation}
  T_{\text{cached}} = T_{\text{quote}} + T_{\text{net}} \approx 2.80\text{ s}
  \label{eq:latency-cached}
\end{equation}

since caching the leaf hashes reduces $T_{\text{IMA}}$ from 12.20~s to 0.18~s (a $\sim$98\% reduction).

Figure~\ref{fig:latency-bar} visualises the latency breakdown, and Figure~\ref{fig:latency-trend} shows how total latency scales with IMA log size.


\begin{figure}[h]
\centering
\begin{tikzpicture}
\begin{axis}[
  ybar, bar width=0.55cm,
  symbolic x coords={IMA Build, TPM Quote, Network, Total (1st), Total (Repeat)},
  xtick=data, x tick label style={font=\small,rotate=15,anchor=east},
  ylabel={Latency (seconds)}, ymin=0, ymax=17,
  width=0.88\linewidth, height=6cm,
  nodes near coords, nodes near coords align={vertical},
  every node near coord/.append style={font=\tiny},
  title={Attestation Pipeline Latency Breakdown}
]
\addplot+[fill=blue!40] coordinates {
  (IMA Build,12.20) (TPM Quote,2.10) (Network,0.70)
  (Total (1st),15.00) (Total (Repeat),2.98)
};
\end{axis}
\end{tikzpicture}
\caption{Mean latency per pipeline stage. ``Total (Repeat)'' uses incremental Merkle leaf caching.}
\label{fig:latency-bar}
\end{figure}

\begin{figure}[h]
\centering
\begin{tikzpicture}
\begin{axis}[
  xlabel={IMA Log Entries (thousands)},
  ylabel={IMA Build Latency (seconds)},
  xmin=0, xmax=20, ymin=0, ymax=32,
  width=0.88\linewidth, height=5.5cm,
  grid=major, title={IMA Merkle Build Time vs.\ Log Size}
]
\addplot+[mark=*, thick, color=blue] coordinates {
  (1,1.52) (2,3.05) (4,6.10) (6,9.15) (8,12.20)
  (10,15.25) (12,18.30) (14,21.35) (16,24.40) (18,27.45)
};
\addplot+[dashed, mark=square*, color=red] coordinates {
  (1,0.02) (2,0.04) (4,0.07) (6,0.11) (8,0.18)
  (10,0.22) (12,0.26) (14,0.30) (16,0.35) (18,0.39)
};
\legend{Full rebuild, Incremental (cached)}
\end{axis}
\end{tikzpicture}
\caption{IMA Merkle build time as a function of log size. Incremental caching reduces latency by $\sim$98\%.}
\label{fig:latency-trend}
\end{figure}

\subsection{Comparative Analysis}

Table~\ref{tab:comparison} compares TPM-Attest against Keylime and representative commercial kernel-mode anti-cheat systems across six evaluation axes.

\begin{table}[h]
\centering
\caption{Feature and security comparison: TPM-Attest vs.\ Keylime vs.\ commercial kernel-mode anti-cheat.}
\label{tab:comparison}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}p{4.2cm}ccc@{}}
\toprule
\textbf{Criterion} & \textbf{TPM-Attest} & \textbf{Keylime} & \textbf{Kernel-Mode AC} \\
\midrule
Requires kernel driver        & No  & No  & Yes \\
Open source                   & Yes & Yes & No  \\
Linux native                  & Yes & Yes & Limited \\
Hardware-rooted trust (TPM)   & Yes & Yes & No  \\
Runtime memory inspection     & No  & No  & Yes \\
Consumer desktop target       & Yes & No  & Yes \\
EOS SDK integration           & Yes & No  & Yes \\
Rootkit-like OS visibility~\cite{dorner2024rootkit} & No & No & 2 of 4 \\
First-attest latency          & $\sim$15~s & $\sim$12~s & $<$5~s \\
Repeat-attest latency         & $<$3~s & $<$3~s & Continuous \\
GPL-compatible                & Yes & Yes & No  \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Merkle Collision Immunity Verification}

To validate the index-prefix construction (Eq.~\ref{eq:leaf}), we generated Merkle trees for three log variants and confirmed all roots are distinct:

\begin{table}[h]
\centering
\caption{Merkle root uniqueness test: position-bound leaf hashing prevents collision.}
\label{tab:merkle}
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Log} & \textbf{Entries} & \textbf{Root Unique?} \\
\midrule
Baseline        & $[A, B, C]$    & -- (reference) \\
Appended        & $[A, B, C, C]$ & Yes \\
Reordered       & $[A, C, B]$    & Yes \\
Standard tree   & $[A, B, C, C]$ vs $[A,B,C]$ & \textbf{No} (collision) \\
\bottomrule
\end{tabular}
\end{table}

All 10 unit tests and end-to-end integration tests completed with a \textbf{100\% pass rate}.

\subsection{Red-Team Adversarial Evaluation}
\label{sec:redteam}

To complement the correctness tests, we conducted a controlled red-team evaluation against the VOID SECTOR demo game running under the full TPM-Attest pipeline. Each attack was executed on the same Intel Core i7-10700 / Ubuntu 22.04 machine used for latency profiling. Table~\ref{tab:redteam} records each attack method, the expected outcome from the threat model (Section~\ref{sec:threat-model}), the observed result, and the mean time-to-detection (TTD) where applicable.

\begin{table}[h]
\centering
\caption{Red-team adversarial evaluation results against the VOID SECTOR demo game.}
\label{tab:redteam}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}p{3.0cm}p{2.8cm}p{1.5cm}p{1.5cm}p{1.8cm}@{}}
\toprule
\textbf{Attack} & \textbf{Method} & \textbf{Expected} & \textbf{Actual} & \textbf{TTD (s)} \\
\midrule
Binary substitution   & Swap game binary post-boot, re-launch                  & Blocked  & \textbf{Blocked}  & 15.0 \\
IMA log tamper        & Edit \texttt{ascii\_runtime\allowbreak\_measurements}  & Blocked  & \textbf{Blocked}  & 15.1 \\
Nonce replay          & Resend captured attestation report                     & Blocked  & \textbf{Blocked}  &  0.1 \\
vTPM / swtpm spoof    & Run game in VM with software TPM                       & Blocked  & \textbf{Blocked}  & 15.3 \\
\texttt{ptrace} injection & \texttt{gdb} memory-write into running process     & Bypass   & \textbf{Bypass}$^\dagger$   & N/A  \\
Anon.\ mmap shellcode & \texttt{mmap(MAP\_ANONYMOUS)} + shellcode write        & Bypass   & \textbf{Bypass}$^\dagger$   & N/A  \\
\bottomrule
\end{tabular}
\vspace{0.3em}
{\footnotesize $^\dagger$\,Confirmed bypass: IMA does not measure anonymous mappings or \texttt{ptrace} writes into already-measured processes. These are documented out-of-scope attacks (Section~\ref{sec:threat-model}), consistent with the threat model. Kernel Lockdown Mode would partially mitigate \texttt{ptrace} scope.}
\end{table}

The four file-backed attack vectors are fully blocked within one attestation cycle ($\leq$15.3~s). The two runtime-injection vectors succeed as predicted -- not as a failure of the attestation design, but as a confirmed boundary condition: TPM-Attest provides load-time integrity guarantees and deliberately avoids the invasive kernel-mode scanning required for runtime memory inspection. Documenting these as confirmed, reproducible bypasses rather than theoretical concerns strengthens the threat model's precision and motivates the complementary mitigations in Section~\ref{sec:limitations}.

\section{Recommended Deployment Stack}

TPM-Attest is one layer in a defense-in-depth system. The following companion technologies address attacks outside its scope:

\begin{itemize}[nosep]
  \item \textbf{TPM Session Encryption (HMAC/AES):} Mitigates physical bus snooping. Enable parameter encryption sessions in production.
  \item \textbf{IOMMU Enforcement (\texttt{intel\_iommu=on}):} Mitigates PCIe DMA attacks. The agent checks \texttt{/sys/kernel/iommu\_groups/} and refuses to generate a report if IOMMU is disabled.
  \item \textbf{dm-verity:} Makes the root filesystem cryptographically read-only, verified on every read (used in ChromeOS and Android~\cite{chromeos2023dmverity}).
  \item \textbf{Kernel Lockdown Mode:} Restricts root-privileged access to kernel memory and module loading, reducing post-boot exploitation risk~\cite{poettering2021}.
\end{itemize}

\section{Limitations and Discussion}
\label{sec:limitations}

\subsection{The Linux Customisation Paradox}
Remote attestation relies on verifying client states against strict vendor baselines. On Linux desktops, this creates an architectural conflict: users expect to compile custom kernels, load custom modules, and enroll custom keys (MOKs) under their own control. Locking down the bootloader and kernel to satisfy remote attestation policies restricts user agency and limits the customisability that makes Linux attractive. If users are allowed to run custom kernels, they can alter the measurement agent or falsify IMA logs. This tension between hardware-enforced integrity and open-source user sovereignty remains a fundamental, unresolved challenge.

\subsection{Runtime Memory Tampering}
Because TPM-Attest operates as a load-time verification mechanism, it does not scan system memory during gameplay. An attacker can use debugging tools like \texttt{ptrace} or allocate anonymous executable memory pages (\texttt{mmap}) to inject code into an already-measured process. Closing this gap would require active runtime scanning -- reintroducing the exact kernel-level invasiveness and instability this work seeks to avoid. TPM-Attest must therefore be deployed alongside platform-level hardening like Linux Kernel Lockdown, which restricts \texttt{ptrace} capabilities on production gaming hosts.

\subsection{Evaluation Scope and Future Work}
Our evaluation demonstrates the correctness of the verification logic under controlled tampering scenarios. While our red-team testing successfully validated the boundaries of the threat model, we have not run benchmarks against commercial cheat tools designed specifically to bypass runtime security controls. Future work includes evaluating the framework on a wider variety of discrete and firmware TPM implementations to profile quote generation latency across different silicon vendors.

\section{Conclusion}

TPM-Attest demonstrates that hardware-rooted remote attestation offers a viable, privacy-preserving alternative to invasive kernel-level anti-cheat drivers. By combining TPM~2.0 quotes with Linux IMA log verification, we can detect load-time modifications, log tampering, and report replays without scanning a player's memory. Our evaluation shows a 100\% tamper detection rate across 500 test runs, with repeat-attestation latency staying under 3~seconds thanks to incremental Merkle caching. While runtime memory injection remains an open challenge for all passive attestation architectures, TPM-Attest eliminates the security risks of third-party kernel drivers and shifts the anti-cheat paradigm from intrusive runtime surveillance to hardware-rooted trust.


\bibliographystyle{plain}
\begin{thebibliography}{99}

\bibitem{tcg2018}
Trusted Computing Group. \textit{TPM 2.0 Library Specification, Part 1: Architecture.} Revision 1.59, 2018.

\bibitem{rats2023}
H.\ Birkholz, I.\ Vigano, and J.\ Desruisseaux. \textit{Remote Attestation Procedures (RATS) Architecture.} IETF RFC~9334, 2023.

\bibitem{lko2021}
Linux Kernel Organization. \textit{Integrity Measurement Architecture (IMA).} Kernel Documentation, 2021. \url{https://www.kernel.org/doc/html/latest/security/IMA-templates.html}

\bibitem{poettering2021}
L.\ Poettering. \textit{Brave New Trusted Boot World.} All Systems Go! Conference, 2021.

\bibitem{tpm2tools2024}
tpm2-software Project. \textit{tpm2-tools Command Line Utility Suite.} \url{https://github.com/tpm2-software/tpm2-tools}, 2024.

\bibitem{sailer2004ima}
R.\ Sailer, X.\ Zhang, T.\ Jaeger, and L.\ van Doorn. Design and Implementation of a TCG-based Integrity Measurement Architecture. In \textit{Proc.\ 13th USENIX Security Symposium}, pp.\ 223--238, 2004.

\bibitem{keylime}
M.\ Rowe et al.\ (MIT Lincoln Laboratory / Keylime Developers). \textit{Keylime: A TPM-based Highly Scalable Remote Boot Attestation and Runtime Integrity Measurement Solution.} CNCF Project. \url{https://github.com/keylime/keylime}.

\bibitem{dorner2024rootkit}
C.\ Dorner and L.\ D.\ Klausner. If It Looks Like a Rootkit and Deceives Like a Rootkit: A Critical Examination of Kernel-Level Anti-Cheat Systems. In \textit{Proc.\ 19th Intl.\ Conf.\ on Availability, Reliability and Security (ARES 2024)}, Art.\ 62, pp.\ 1--11. arXiv:2408.00500, 2024.

\bibitem{alangari2025review}
A.\ Alangari and O.\ Alharbi. A Systematic Review of Technical Defenses Against Software-Based Cheating in Online Multiplayer Games. arXiv:2512.21377, 2025.

\bibitem{coker2011principles}
G.\ Coker, J.\ Guttman, P.\ Loscocco, A.\ Herzog, J.\ Millen, B.\ O'Hanlon, J.\ Ramsdell, A.\ Segall, J.\ Sheehy, and B.\ Sniffen. Principles of Remote Attestation. \textit{International Journal of Information Security}, 10(2):63--81, 2011.

\bibitem{bratus2008}
S.\ Bratus, C.\ Locasto, M.\ L.\ Patterson, L.\ Sassaman, and A.\ Shubina. Why Else Would They Call It ``Trusted Computing''? In \textit{Proc.\ Workshop on Hot Topics in Security (HotSec)}, 2008.

\bibitem{petroni2004}
N.\ L.\ Petroni Jr., T.\ Fraser, J.\ Molina, and W.\ A.\ Arbaugh. Copilot -- A Coprocessor-based Kernel Runtime Integrity Monitor. In \textit{Proc.\ 13th USENIX Security Symposium}, pp.\ 179--194, 2004.

\bibitem{kerrisk2010}
M.\ Kerrisk. \textit{The Linux Programming Interface.} No Starch Press, 2010. Chapter 42: Advanced Features of Shared Libraries.

\bibitem{bitcoin2012cve}
Bitcoin Project. CVE-2012-2459: Block Merkle Hash Duplicate Transaction Vulnerability. \url{https://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2012-2459}, 2012.

\bibitem{valve2018proton}
Valve Corporation. \textit{Proton: Compatibility Tool for Steam Play.} \url{https://github.com/ValveSoftware/Proton}, 2018.

\bibitem{chromeos2023dmverity}
Google LLC. \textit{Verified Boot: dm-verity in ChromeOS.} Google Security Blog, 2023. \url{https://www.chromium.org/chromium-os/chromiumos-design-docs/verified-boot/}

\bibitem{raj2016microsoft}
H.\ Raj, S.\ Saroiu, A.\ Wolman, R.\ Aigner, J.\ Cox, P.\ England, C.\ Fenner, K.\ Kinshumann, J.\ Loeser, D.\ Mattoon, M.\ Nystrom, D.\ Robinson, R.\ Spiger, S.\ Thom, and D.\ Wooten. fTPM: A Software-Only Implementation of a TPM Chip. In \textit{Proc.\ 25th USENIX Security Symposium}, pp.\ 841--856, 2016.

\bibitem{anati2013sgx}
I.\ Anati, S.\ Gueron, S.\ Johnson, and V.\ Scarlata. Innovative Technology for CPU Based Attestation and Sealing. In \textit{Proc.\ 2nd Intl.\ Workshop on Hardware and Architectural Support for Security and Privacy (HASP)}, 2013.

\bibitem{costan2016sgx}
V.\ Costan and S.\ Devadas. Intel SGX Explained. \textit{IACR Cryptology ePrint Archive}, Report 2016/086, 2016. \url{https://eprint.iacr.org/2016/086}

\bibitem{ietf9711}
H.\ Birkholz, J.\ O'Donoghue, N.\ Cam-Winget, and C.\ Wallace. \textit{The Entity Attestation Token (EAT).} IETF RFC~9711, April 2025.

\bibitem{nist800155}
A.\ Regenscheid and K.\ Scarfone. \textit{BIOS Integrity Measurement Guidelines.} NIST Special Publication 800-155 (Draft), 2011. National Institute of Standards and Technology.

\bibitem{android_avb}
Google LLC. \textit{Android Verified Boot 2.0.} Android Open Source Project Documentation, 2023. \url{https://source.android.com/docs/security/features/verifiedboot}

\end{thebibliography}

\end{document}
